---
title: "simulatedLineup2"
format: html
editor: visual
---

```{r}
library(tidyverse)
df <- read_csv("savantData21_24.csv")
savant_data_2025 <- read_csv("savantData25.csv")
```

Filter the savant data to only get the relevant columns for finding the outcomes and the base-out matrices
```{r}
filterSavantData <- function(df){
  filtered_df <- df |> 
    select(game_id, inning, half_inning, outs, pre_runner_1b_id, pre_runner_2b_id, pre_runner_3b_id,
           batter_id, bat_side, pitcher_id, pitch_hand, events, home_score, away_score, post_home_score,
           post_away_score, at_bat_number) |> 
    filter(!is.na(events)) |>  
    filter(inning < 10) |> 
    arrange(game_id, inning, half_inning)
  return(filtered_df)
}

```


Find the percent outcomes for all 12 outcomes. The 6 outcomes I chose for both left and right handed pitching
```{r}

createBatterOutcomes <- function(df){
  batterOutcomes <- df |> 
    filter(!events %in% c("truncated_pa", "sac_bunt", "sac_fly", "sac_fly_double_play",
                          "sac_bunt_double_play", "catcher_interf")) |> 
    group_by(batter_id) |> 
    summarize(
      n_PA_L = sum(pitch_hand == "L"),
      p_BB_L = sum((pitch_hand == "L" & (events %in% c("walk", "hit_by_pitch")))) / n_PA_L,
      p_1B_L = sum((pitch_hand == "L" & events == "single")) / n_PA_L,
      p_2B_L = sum((pitch_hand == "L" & events == "double")) / n_PA_L,
      p_3B_L = sum((pitch_hand == "L" & events == "triple")) / n_PA_L,
      p_HR_L = sum((pitch_hand == "L" & events == "home_run")) / n_PA_L,
      p_out_L = 1 - p_BB_L - p_1B_L - p_2B_L - p_3B_L - p_HR_L,
      n_PA_R = sum(pitch_hand == "R"),
      p_BB_R = sum((pitch_hand == "R" & (events %in% c("walk", "hit_by_pitch")))) / n_PA_R,
      p_1B_R = sum((pitch_hand == "R" & events == "single")) / n_PA_R,
      p_2B_R = sum((pitch_hand == "R" & events == "double")) / n_PA_R,
      p_3B_R = sum((pitch_hand == "R" & events == "triple")) / n_PA_R,
      p_HR_R = sum((pitch_hand == "R" & events == "home_run")) / n_PA_R,
      p_out_R = 1 - p_BB_R - p_1B_R - p_2B_R - p_3B_R - p_HR_R,
    )
  return(batterOutcomes)
}

```

Create easier to handle on base flags and add runs scored as a column
```{r}
filtered_df <- filterSavantData(df)
filtered_df <- filtered_df |> 
  mutate(
        runner_1B = if_else(pre_runner_1b_id > 0, 1, 0, missing=0),
        runner_2B = if_else(pre_runner_2b_id > 0, 1, 0, missing=0),
        runner_3B = if_else(pre_runner_3b_id > 0, 1, 0, missing=0)
      ) |> 
      select(-pre_runner_1b_id, -pre_runner_2b_id, -pre_runner_3b_id) |> 
  mutate(
    runsScored = if_else(half_inning == "top", post_away_score - away_score, post_home_score - home_score)
  ) |> 
  select(-post_away_score, -post_home_score, -away_score, -home_score) |> 
  group_by(game_id) |> 
  arrange(game_id, at_bat_number) |> 
  mutate(
    runsScoredLag = lag(runsScored)
  ) |> 
  ungroup()

```


Create the matrix for all input codes and output codes
```{r}
createCode <- function(df, inputFlag){
  df2 <- df |> 
    mutate(
      outcomeID = case_when(
        events %in% c("walk", "hit_by_pitch", "catcher_interf") ~ 0,
        events == "single"                    ~ 1,
        events == "double"                    ~ 2,
        events == "triple"                    ~ 3,
        events == "home_run"                  ~ 4,
        TRUE                                  ~ 5 
      )
    )
  
  if(inputFlag) {
    df3 <- df2 |> mutate(
      rawCode = 10000 * runner_1B + 1000 * runner_2B + 100 * runner_3B + 10* outs + outcomeID
    )
  } else{
    df3 <- df2 |>  mutate(
      rawCode = 10000 * runner_1B + 1000 * runner_2B + 100 * runner_3B + 10* outs + runsScoredLag
    )
  }
  df3 |> 
    mutate(code = sprintf("%05d", rawCode)) |> 
    pull(code)
}
```


Generate all possible input and output code values
```{r}
full_inputs <- expand.grid(
    r1 = 0:1, r2 = 0:1, r3 = 0:1,
    outs = 0:2, outcomeID = 0:5
  ) |> 
  transmute(
    code = sprintf(
      "%05d",
      10000*r1 + 1000*r2 + 100*r3 + 10*outs + outcomeID
    )
  ) |> 
  pull(code)

full_outputs <- expand.grid(
    r1 = 0:1, r2 = 0:1, r3 = 0:1,
    outs = 0:2, runsScored = 0:4
  ) |> 
  transmute(
    code = sprintf(
      "%05d",
      10000*r1 + 1000*r2 + 100*r3 + 10*outs + runsScored
    )
  ) |> 
  pull(code)
```

Tag each plate appearance with its input and output code
```{r}
df_codes <- filtered_df |> 
  mutate(
    inputCode = createCode(filtered_df, TRUE),
    outputCode = createCode(filtered_df, FALSE)
  ) |> 
  group_by(game_id) |> 
  arrange(at_bat_number, .by_group = TRUE) |> 
  mutate(
    nextOutputCode = lead(outputCode)
  ) |> 
  ungroup()

#Removes the last plate appearance in a game since it has no valid output code
df_codes <- df_codes |> 
  filter(!is.na(nextOutputCode))

```

Count and turn into row-conditional probabilities
```{r}
prob_table <- df_codes |> 
  count(inputCode, nextOutputCode) |> 
  complete(
    inputCode = full_inputs,
    nextOutputCode = full_outputs,
    fill = list(n = 0)
  ) |> 
  group_by(inputCode) |> 
  mutate(
    p = n / sum(n)
  ) |> 
  ungroup()
```


Pivot into a 144x120 matrix
```{r}
prob_matrix <- prob_table |> 
  select(inputCode, nextOutputCode, p) |> 
  pivot_wider(
    names_from = nextOutputCode,
    values_from = p
  ) |> 
  arrange(inputCode) |> 
  column_to_rownames("inputCode")

prob_matrix <- prob_matrix |>  rownames_to_column(var = "inputCode")

```


Test code to find edge cases
```{r}
df_codes |> 
  filter(inputCode == "00000", nextOutputCode == "10010")

test <- df_codes |> 
  filter(game_id == 717644, , inning == 8)

test <- df |> 
  filter(game_id == 717644, inning == 8)

test <- df_codes |> 
  group_by(game_id) |> 
  mutate(
    runsScoredLag = lag(runsScored)
  )

```


Simulate an outcome from a Plate Appearance
```{r}
simulateOutcome <- function(batterID, pitcher_hand, outcomesTable){
  batterOutcomes <- outcomesTable |> 
    filter(batter_id == batterID)
    
  if(pitcher_hand == "L"){
      probs <- c(
        BB = batterOutcomes$p_BB_L,
        Single = batterOutcomes$p_1B_L,
        Double = batterOutcomes$p_2B_L,
        Triple = batterOutcomes$p_3B_L,
        Home_Run = batterOutcomes$p_HR_L,
        Out = batterOutcomes$p_out_L
      )
  }
  if(pitcher_hand == "R"){
      probs <- c(
        BB = batterOutcomes$p_BB_R,
        Single = batterOutcomes$p_1B_R,
        Double = batterOutcomes$p_2B_R,
        Triple = batterOutcomes$p_3B_R,
        Home_Run = batterOutcomes$p_HR_R,
        Out = batterOutcomes$p_out_R
      )
  }
  
  outcome <- sample(names(probs), size = 1, prob = probs)
  if(outcome == "BB"){
    outcome <- 0
  }
  if(outcome == "Single"){
    outcome <- 1
  }
  if(outcome == "Double"){
    outcome <- 2
  }
  if(outcome == "Triple"){
    outcome <- 3
  }
  if(outcome == "Home_Run"){
    outcome <- 4
  }
  if(outcome == "Out"){
    outcome <- 5
  }
  return(outcome)
}

batterOutcomes2025 <- filterSavantData(savant_data_2025) |> 
  createBatterOutcomes()

simulateOutcome(656941, "L", batterOutcomes2025)
```
Function to get output code from input code
```{r}
getOutputCode <- function(input, probMatrix){
  selectedInputCode <- probMatrix |> 
    filter(inputCode == input)

  probs <- selectedInputCode |> 
    select(-inputCode) |> 
    unlist(use.names = FALSE)
  #print(probs)
  
  codes <- names(selectedInputCode)[-1]
  #print(codes)

  outputCode <- sample(codes, size = 1, prob = probs)
  return(outputCode)
}

```

Simulate an Inning
```{r}
simulateInning <- function(lineup, lineupPosition, pitcherHand){
  lastOutputCode <- "00000"
  outs <- 0
  runs <- 0
  
  repeat{
    inputCode <- lastOutputCode
    outcome <- simulateOutcome(lineup[lineupPosition], pitcherHand, batterOutcomes2025)
    outcome <- as.character(outcome)
    substr(inputCode, 5, 5) <- outcome #Change the last digit of the input code
    outputCode <- getOutputCode(inputCode, prob_matrix)
    
    runsAdded <- as.integer(substr(outputCode, 5, 5))
    outs <- as.integer(substr(outputCode, 4, 4))
    
    runs <- runs + runsAdded
    
    #Prints out every at bat
#    cat(sprintf(
#      "PA #%2d | Batter: %s | Outcome: %s | +Outs: %d | +Runs: %d | TotRuns: %d\n, | InputCode: %s\n",
#      lineupPosition,
#      as.character(lineup[lineupPosition]),
#      outcome,
#      outs,
#      runsAdded,
#      runs,
#      inputCode
#    ))
    
    #prep for next batter
    lastOutputCode <- outputCode
    lineupPosition <- ((lineupPosition %% length(lineup)) + 1)
    
    #Break when it goes back to "00000" (Start of inning state)
    if (lastOutputCode == "00000") break
  }
  return(list(runs = runs, lineupPosition = lineupPosition))
  
}

testLineup <- c(656941, 656941, 656941, 656941, 656941, 656941, 656941, 656941, 656941)
test <- simulateInning(testLineup, 1, "L")

```

Simulate Game
```{r}
simulateGame <- function(lineup, pitcherHand){
  lineupPosition <- 1
  totalRuns <- 0
  
  for(i in 1:9){
    inningOuput <- simulateInning(lineup, lineupPosition, "L")
    
    totalRuns <- totalRuns + inningOuput$runs
    lineupPosition <- inningOuput$lineupPosition
  }
  return(totalRuns)
}

testLineup <- c(680776, 680776, 680776, 680776, 680776, 680776, 680776, 680776, 680776)
simulateGame(testLineup, "L")
```

```{r}



```






